Data Appendix to \"The impact of weather events on schooling outcomes\

### 2020-03-25 data_appendix.Rmd
adding to the data patterns part

```{r set dfSummary css style, echo = F, include = F}
st_css()
```
# Appendix description
*Your Data Appendix should begin with a brief statement explaining its purpose like the following one.*

@@

- **hh_id:** The household ID, which is the same for each member of the household. It is made up of letters and numbers. 
- **nro1:** The number of the person in each household.
- **assist:** Binary variable, whether or not the individual assists school 
- **urb_rur:** Binary variable, whether the household is in an urban or rural area
- **reason:** the reason why the individual is not assisting school if they are not 
- **members:** number of household members excluding servants 
- **sex:** the sex of the individual, Male = 1, Female =2 
- **age:** How old the individual is, in years 
- **relation:** The relationship of the individual to the head of household 
- **literacy:** Binary variable, whether a person can read and write 
- **register:** Binary variable, whether an individual registered for school or not 
- **grade:** For what grade is the person registered 
- **repeated:** whether the person has repeated the grade they are currently registered in 
- **dept:** the department in which the person lives

@@

```{r import household data}
data_2005 <- read_dta(file.path("Raw data","seminar_paper_data","2005_hhsurvey.dta")) %>% 
  rename(hh_id = folio,
        assist = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy = s3_01,
        register = s3_04,
        grade = s3_05b,
        repeated = s3_07)

hh_2005 <- data_2005 %>% 
  select(hh_id, assist, reason, members, sex, age,
         relation, literacy, register, grade, repeated, depto, urb_rur) %>% 
  mutate(year = 2005,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"))
```

@@

Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales

**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
**Date Downloaded:** Identify when you downloaded the dataset.  
**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
**Unit of observation:** What distinguishes different rows in your dataset?  
**Dates covered:** What time frame does the data cover?

@@

Describe in a step-by-step fashion how an interested user could obtain the data.

### Importable version (if necessary)

**Filename(s):** importable-data/filename-importable.csv

In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.

@@

This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225. It is shown on a map with different layers.

### Importable version (if necessary)

**Filename(s):** importable-data/filename-importable.csv

In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.

@@

# Data Processing and Combination
To create a complete dataset, I merged the household data with the floods data and the droughts data. Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis.
*This section should include a discussion of the processing and merging steps needed to create your basic data. The code to implement these steps should be included in chunks in this section. Once the final merged data has been created, you should use the dfSummary function again to summarize the data you will be using. You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.*

```{r merging datasets}
#I would recommend combining all the years of household data and then merging with the weather data

total_2005 <- hh_2005 %>% 
  left_join(flood_summary) %>%
  left_join(drought_summary)
```

@@

This section should include a description of all the variables that are used in your final analysis. At the end of the section, you should save all of these variables in the processed_data folder of your repository.

@@

In the household data, the sample of children in 2005 was

*This section should include a discussion of any data patterns you notice based on the summaries created in the code above.*

>>>>

```{r set dfSummary css style, echo = F, include = F}
st_css()
```
# Appendix description
This data appendix details the steps of collecting, processing, and merging the data needed to create the complete data set showing information about schooling and weather events. It concludes with a discussion of patterns found in the data.

@@

- **hh_id:** The household ID, which is the same for each member of the household. It is made up of letters and numbers. 
- **nro1:** The number of the person in each household.
- **assist:** Binary variable, whether or not the individual assists school 
- **urb_rur:** Binary variable, whether the household is in an urban or rural area
- **reason:** the reason why the individual is not assisting school if they are not 
- **members:** number of household members excluding servants 
- **sex:** the sex of the individual, Male = 1, Female =2 
- **age:** How old the individual is, in years 
- **relation:** The relationship of the individual to the head of household 
- **literacy:** Binary variable, whether a person can read and write 
- **register:** Binary variable, whether an individual registered for school or not 
- **grade:** For what grade is the person registered 
- **repeated:** whether the person has repeated the grade they are currently registered in 
- **dept:** the department in which the person lives 
- **income:** the per capita monthly income for each household, in bolivianos

@@

```{r import household data}
data_2005 <- read_dta(file.path("Raw data","seminar_paper_data","2005_hhsurvey.dta")) %>% 
  rename(hh_id = folio,
        assist = s3_08,  
        reason = s3_09, 
        members = nm_y,
        sex = s1_02,
        age = s1_03,
        relation = s1_05,
        literacy = s3_01,
        register = s3_04,
        grade = s3_05b,
        repeated = s3_07,
        income = y_perc1A)

hh_2005 <- data_2005 %>% 
  select(hh_id, assist, reason, members, sex, age,
         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
  mutate(year = 2005,
         dept = case_when(depto == "1" ~ "Chuquisaca",
                          depto == "2" ~ "La Paz",
                          depto == "3" ~ "Cochabamba",
                          depto == "4" ~ "Oruro",
                          depto == "5" ~ "Potosi",
                          depto == "6" ~ "Tarija",
                          depto == "7" ~ "Santa Cruz",
                          depto == "8" ~ "Beni",
                          depto == "9" ~ "Pando"))
```

@@

**Date Downloaded:** October 2019
**Filename(s):** raw_data/filename.csv 
**Unit of observation:** Municipality  
**Dates covered:** 2002-2012

@@

This data can be found at the following website: http://geosinager.defensacivil.gob.bo/maps/226. It is associated with a map showing the flood occurrences information spatially, and to get the data in a non-spatial format, the flood layer should be downloaded. It may be downloaded in a variety of different formats, including csv.

@@

This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225. It is shown on a map with different layers, and the raw data may be downloaded in a variety of different forms, including csv.

@@

# Data Processing and Combination

To create a complete dataset, I merged the household data with the floods data and the droughts data. Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis in each file. I renamed several variables in the household survey data set to make them more understandable. I also created the year variable for the household data to ensure that it matched up correctly with the weather data. For the flood and drought data, I grouped the occurrences by department instead of by municipality so they would match the household data, and I made the department character numeric for the merge.

*This section should include a discussion of the processing and merging steps needed to create your basic data. The code to implement these steps should be included in chunks in this section. Once the final merged data has been created, you should use the dfSummary function again to summarize the data you will be using. You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.*

``` {r merging datasets}
total_2005 <- hh_2005 %>% 
  left_join(flood_summary) %>%
  left_join(drought_summary)

export_summary_table(dfSummary(total_2005))

total_2005 <- readRDS(file = analysis_variables.rds")

@@

The variables used in the final analysis are the variables from the flood and drought summaries, showing the number of occurrences in each department per year. The other variables used are from the household data, showing demographic information as well as the dependent variables related to school assistance and registration. The variables "assist" and "register" tell whether the child is currently registered and attending school, and the variables "grade" and "literacy" give further educational information showing what grade the child is in and whether they can read and write. The variables "urb_rur" and "dept" give geographic information, indicating whether the child's household is located in an urban area or a rural area and in what department of the country they are living. Demographic variables such as age, sex, relation to the head of household, number of household members, and income can be used in the analysis as controls.

At the end of the section, you should save all of these variables in the processed_data folder of your repository.

@@

The "assist" variable shows that around 60% of people who are registered for school currently assist, while 40% do not assist. The "reason" variable shows that two most common reasons for missing school are for work (31.3%) or vacation (20.5). Out of the people who are registered for school, 6.6% are repeating the grade they are currently in.

The patterns for households show that the average household has 5.3 members, and the average per capita monthly income per household is 258 bolivianos. Over half of the households in the sample are located in urban areas (56.3), while under half are located in rural areas (43.7). La Paz, Santa Cruz, and Cochabamba are the three departments with the most households in this sample, which corresponds to their overall population distribution.

The patterns in the weather data show that every department experienced multiple floods in 2005, while many but not all experienced droughts. The place that experienced the highest number of flood occurrences overall was Cochabamba, while Beni experienced 13 droughts in 2005.
